A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals

### 2019-10-09 paper/paper.experiments.tex
Update paper.experiments.tex
SM reviewed experiments - not finished yet

% JK: I think give an overview here of the experiments here
%     and what you're going to show in this section (and why)
In this section we compare the schedules created by solving the ILP formulation
given in Section~\ref{sec:methods} to schedules that were manually generated,
in terms of adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
We also want to analyze the efficiency of the ILP approach in generating schedules
by its runtime on a variety of instances that may be found in the real-world.
Our goal is to show that the schedules provided by solving the ILP successfully
enforce all hard constraints and improve satisfiability of soft constraints, 
as well as to demonstrate that our ILP formulation can be used for a wide range
of configurations.

@@

First, we evaluated the ILP solution by comparing it with the
% JK: I think it would be good to give a short name for manual and the proposed
%     data -- e.g. LP Solution as in Table 4 and Manual Generation.
%     Then check for consistency throughout.
%     While lack of synonyms can sound repetitive, it can help avoid confusion.
manual generation, as in Table~\ref{tbl:2018-schedule-comparison}. Specifically, we examined the adherence of each
schedule to the constraints presented in Table~\ref{tbl:constraint-summary}. As
shown in Table~\ref{tbl:constraints-comparison}, the ILP solution satisfied all 
hard constraints. In contrast,
% JK: "mandatory" = "hard"? if possible just use one.
manual generation did not satisfy all hard constraints. In
particular, we see that the manual generation assigned clinicians to multiple
consecutive blocks in all four years. Moreover, the manual generation did not have
an equal distribution of weekends and holidays for all four years of data.
Considering all objectives, we see that the ILP solution outperforms manual generation
in all four years, by accommodating almost all time-off requests and
ensuring that weekends are always assigned close to blocks.

@@

% JK: Introduce what you did first and then the figure showing results (see above)
The effect of increasing the number of clinicians and number of services %divisions %do you mean
%services?
on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
We executed the algorithm for
$S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in
total across all services. 
In a department providing a single service, increasing the number of clinicians
did not affect the runtime, and we were able to find an ILP solution in all four cases
within 1 second.
For 2 concurrent services, a roster of 30 or more
clinicians becomes impractical to schedule, as searching for a solution required
over 24 hours. We saw similar issues for a roster of 20 or more clinicians
assigned to a division with 3 concurrent services. However, when removing the
NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3
services, and we were able to generate a schedule with upwards of 50 clinicians
in under 1.5 seconds.  %what does division mean? do you mean services?

>>>>

% JK: I think give an overview here of the experiments here
%     and what you're going to show in this section (and why)
% check verb tense - use past tense I think given flow
Our goal was to determine if the schedules provided by solving the ILP 
could successfully
enforce all hard constraints; improve satisfiability of soft constraints compared with ??; %SM: found "improve satisfiability" a bit hard to follow. revise this phrase for clarity
and assess if our ILP formulation can be used for a wide range
of configurations.  %list objectives with numbers (i, ii, iii) or separated by semi-colons - helps the reader re: flow
First, we compared the schedules created by solving the ILP formulation
given in Section~\ref{sec:methods} to schedules that were manually generated,
with respect to adherence to the hard and soft constraints outlined in Section~\ref{sec:problem}.
We then examined the efficiency of the ILP approach in generating schedules
by its runtime on a variety of instances that may be found in the real-world.

@@

First, we evaluated the ILP solution by comparing it with the
% JK: I think it would be good to give a short name for manual and the proposed
%     data -- e.g. LP Solution as in Table 4 and Manual Generation.		%SM: great suggestion!
%     Then check for consistency throughout.
%     While lack of synonyms can sound repetitive, it can help avoid confusion.  %SM: agree
manual generation, as in Table~\ref{tbl:2018-schedule-comparison}. Specifically, we examined the adherence of each
schedule to the constraints presented in Table~\ref{tbl:constraint-summary}. As
shown in Table~\ref{tbl:constraints-comparison}, the ILP solution satisfied all 
hard constraints. In contrast,
% JK: "mandatory" = "hard"? if possible just use one.
manual generation did not satisfy all hard constraints. In
particular, we see that the manual generation assigned clinicians to multiple
consecutive blocks in all four years. Moreover, the manual generation did not have
an equal distribution of weekends and holidays for all four years of data.
Considering all objectives, we see that the ILP solution outperforms manual generation
in all four years, by accommodating almost all time-off requests and
ensuring that weekends are always assigned close to blocks.

@@

% JK: Introduce what you did first and then the figure showing results (see above)
The effect of increasing the number of clinicians and number of services %divisions %do you mean
%services?
on the runtime of the program is shown in Table~\ref{tbl:runtime-services-clinicians-comparison}.
We executed the algorithm for
$S = \{1, 2, 3\}$ total services and $C = \{10, 20, 30, 50\}$ clinicians in
total across all services. 
In a department providing a single service, increasing the number of clinicians
did not affect the runtime, and we were able to find an ILP solution in all four cases
within 1 second.
For 2 concurrent services, a roster of 30 or more
clinicians becomes impractical to schedule, as searching for a solution required
over 24 hours. We saw similar issues for a roster of 20 or more clinicians
assigned to a division with 3 concurrent services. However, when removing the
NCB constraint, we saw a great improvement in runtime for divisions with 2 and 3
services, and we were able to generate a schedule with upwards of 50 clinicians
in under 1.5 seconds.  %what does division mean? do you mean services?  %SM: make sure division vs. services defined in introduction, but may need to repeat again in problem statement to help the reader
